Conservation decision-making under long transients

### 2019-07-26 notebook/ghost-fit.Rmd
add may outbreak model
slides WIP

lows <- means %>% dplyr::slice(1:10) %>% pull(reps)
highs <- means %>% dplyr::slice((n()-10):n()) %>% pull(reps)
coding <- function(reps) case_when(
  reps %in% lows ~ "low",
  reps %in% highs ~ "high"
)
df <- data %>% mutate(coding = coding(reps)) %>% na.omit()
df %>% ggplot(aes(t, x, group=reps, color = coding)) + geom_line(alpha=0.5)

@@

@@

@@

>>>>

lows <- means %>% dplyr::slice(1:10) %>% pull(reps)
highs <- means %>% dplyr::slice((n()-10):n()) %>% pull(reps)
coding <- function(reps) case_when(
  reps %in% lows ~ "low",
  reps %in% highs ~ "high"
)
extremes <- data %>% mutate(coding = coding(reps)) %>% na.omit()

write_csv(extremes, "../data/extremes.csv")
extremes %>% ggplot(aes(t, x, group=reps, color = coding)) + geom_line(alpha=0.5)

@@

----

@@

```{r}
mean <- x_t + r * x_t * (1 - x_t / K) - a * x_t ^ Q / (x_t ^ Q + H ^ Q)
distribution(x_t1) <- normal(mean, sigma * x_t)
m <- model(a)
```

```{r }
system.time({
  draws2 <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws2)
```

```{r}
samples2 <-  
  map_dfr(draws2, 
          function(x) data.frame(x, t = 1:dim(x)[1]), 
          .id = "chain") %>% 
  gather(variable, value, -t, -chain)

samples2 %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

## Random sample

```{r}

random <- sample(unique(data$reps), 1)
wide <- data %>% 
  filter(reps %in% random) %>%
  spread(reps, x) %>%
  select(-t) %>% 
  as.matrix()
n <- dim(wide)[1]

@@

```{r }
system.time({
  draws3 <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws3)
```

```{r}
samples3 <-  
  map_dfr(draws3, 
          function(x) data.frame(x, t = 1:dim(x)[1]), 
          .id = "chain") %>% 
  gather(variable, value, -t, -chain)

samples3 %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

```{r}
```

### 2019-07-26 notebook/ghost-sims.Rmd
slides WIP

```{r message = FALSE}
library(tidyverse)
library(furrr)
```

@@

```{r}
theory <- 
  function(p){
    tibble(x= seq(0,2, length.out = 100)) %>%
    mutate(g = growth(x,p), 
           c = consumption(x,p)) %>%
    mutate(potential = - cumsum(g - c)) %>%
    gather(curve, y, -x, -potential) 
}
```

@@

stochastic_sim <- function(f, p){
  x <- numeric(p$N)
  x[1] <- p$x0
  dBt <- rnorm(p$N, 0, p$sigma)
  for(t in 1:(p$N-1)){
    x[t+1] <- max( 
                  f(x[t], p) + x[t] * dBt[t],  # arguably x[t] * dB[t] would be a better model!
                  0)
  }
  data.frame(t = 1:p$N, x)
}

@@

>>>>

```{r message = FALSE}
library(tidyverse)
library(furrr)
set.seed(12345)
```

@@

```{r}
theory <- 
  function(p){
    tibble(x= seq(0,2, length.out = 100)) %>%
    mutate(growth = growth(x,p), 
           consumption = consumption(x,p)) %>%
    mutate(potential = - cumsum(g - c)) %>%
    gather(curve, y, -x, -potential) 
}
```

@@

stochastic_sim <- function(f, p){
  x <- numeric(p$N)
  x[1] <- p$x0
  dBt <- rnorm(p$N, 0, p$sigma)
  for(t in 1:(p$N-1)){
    x[t+1] <- max( 
                  f(x[t], p) + x[t] * dBt[t],
                  0)
  }
  data.frame(t = 1:p$N, x)
}

@@

```{r}
det <- det_sim(may,p) 
det %>% ggplot(aes(t, x)) + geom_line() + theme_bw()
```

### 2019-07-26 slides/boettiger.Rmd
slides WIP

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "custom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

@@

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .6), rgba(150, 150, 150, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, middle, inverse

@@

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .6), rgba(150, 150, 150, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, middle, inverse

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-size: 120%
class: center, middle, inverse

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-size: 120%
class: center, top, inverse

@@

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
class: center, top, inverse

>>>>

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev.args=list(bg="transparent"), 
                      echo = FALSE, message=FALSE, warning=FALSE,
                      fig.width=11, fig.height=6.5, cache = TRUE)

library(tidyverse)
library(ggthemes)

theme_set(
 theme_bw(base_size=26) + 
  theme(line = element_line(colour = "white", size = 1.5),
        rect = element_rect(fill = "transparent", colour = "white"),
        text = element_text(colour = "white"),
        axis.text = element_text(colour = "white"),
        axis.ticks = element_line(colour = "white", size = 1.5),
        panel.border = element_rect(colour = "white", size = 1.5),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_blank(), 
        legend.background = element_blank(), 
        legend.key = element_rect(fill = "transparent", colour = "transparent"), 
        strip.background = element_blank(),
        strip.text = element_text(colour = "white"),
        panel.grid = element_blank()))

source("../R/ghost.R")
```

@@

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .1), rgba(150, 150, 150, .4)), url(img/city-mit-license.jpg)
background-size: 140%
class: center, top, inverse

@@

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .1), rgba(150, 150, 150, .4)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, top, inverse

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-position: center
background-size: 120%
class: center, top, inverse

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-position: center
background-size: 120%
class: center, top, inverse

@@

<br/>
Hastings et al. *Science* (2018)

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
background-position: center
class: left, top, inverse

# Examples

Outbreak: disease, pest, invasive species

- Low endemic level
- outbreak
- saturation

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
background-position: center
class: center, top, inverse

# A look at the model

$$X_{t+1} = X_t + \underbrace{X_t r \left(1 -\frac{X_t}{K} \right)}_{\textrm{logistic growth}}  - \underbrace{\frac{a X_t ^ Q}{X_t^ Q + H ^ Q}}_{\textrm{saturating consumption}} + \underbrace{\xi_t X_t}_{\textrm{environmental noise}}$$

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 120%
background-position: center
class: center, top, inverse

## Potential well diagram

```{r echo=FALSE}
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = .02, a=0.023, 
          N = 4e3, x0 = 0.2, N = 1e4)
theory(p) %>%
  ggplot(aes(x, y, col=curve)) +
  geom_line(lwd = 2)

```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 120%
background-position: center
class: center, top, inverse

## Potential well diagram

```{r echo=FALSE}
theory(p) %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd = 2, col="white")

```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 120%
background-position: center
class: center, top, inverse

## Dynamics from the deterministic skeleton

```{r echo=FALSE}
det <- det_sim(may,p) %>% mutate(reps=1) %>% filter(t < 3000)
det  %>%
  ggplot(aes(t, x)) + 
  geom_line(lwd = 2, col="white")
```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Under the stochastic model

```{r}
extremes <- read_csv("../data/extremes.csv")
#extremes %>%  filter(coding == "high", t < 3000) %>% pull(reps) %>% unique()

extremes %>% 
  filter(coding == "high", reps == 14, t < 3000) %>%  # %>% pull(reps) %>% unique()
  ggplot(aes(t, x)) + 
  geom_line(lwd = 2, col="white")

```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Under the stochastic model

```{r echo=FALSE}
#extremes %>%  filter(coding == "low", t < 3000) %>% pull(reps) %>% unique()
extremes %>% 
  filter(coding == "low", reps == 47, t < 3000) %>%
  ggplot(aes(t, x)) + 
  geom_line(lwd = 2, col="white")

```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .95)), url(img/royalsociety/gentoo-penguin.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Under the stochastic model

```{r}
data <- vroom::vroom("../data/reps.csv") %>% filter(t < 3000)
mean <- data %>% group_by(t) %>% summarise(x = mean(x)) %>% mutate(reps = 1)
```

```{r}
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white")
```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .95)), url(img/royalsociety/gentoo-penguin.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Under the stochastic model

```{r echo=FALSE}
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white") + 
  geom_line(data = det, lwd=2, color = "white")
```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .95)), url(img/royalsociety/gentoo-penguin.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Under the stochastic model

```{r echo=FALSE}
data %>% 
  ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white") + 
  geom_line(data = mean, lwd=2, color = rgb(.6, .6, 1))
```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
class: center, top, inverse
background-position: center

# Challenges to Model Inference

@@
